---
title: "Extract Transform Load"
author: "Karen Jiang"
date: "2/26/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(readr)
```

## Understanding the Data

### Exploratory Data Analysis

We are examining 3 datasets from CMS.

* ACO general information
* Beneficiaries information by ACO
* Fee-for-Service Expenditures by ACO

To better understand how MSSP scoring and classifications for ACOs, information can be found in the  [MSSP Shared Savings and Losses Assignement and Methodology ]("https://www.cms.gov/Medicare/Medicare-Fee-for-Service-Payment/sharedsavingsprogram/Downloads/Shared-Savings-Losses-Assignment-Spec-V7.pdf").


```{r read_data, message=FALSE}

# Read in .csv data

aco <- 
  read.csv("data/2018_Shared_Savings_Program__SSP__Accountable_Care_Organizations__ACO__PUF.csv") %>%
  select(-contains("N_"), -contains("P_")) 

beneficiaries <- 
  read_csv("data/2018_Number_of_ACO_Assigned_Beneficiaries_by_County_PUF.csv")

ffs <- 
  read_csv("data/2018_County-level_FFS_Data_for_Shared_Savings_Program_Benchmark_PUF.csv")
```


### Exploratory Data Analysis
```{r aco, echo=FALSE}

aco %>%
  count(Agree_Type)


tracks <- aco %>%
  summarise(
    sum(Initial_Track_1),
    sum(Initial_Track_1_Plus),
    sum(Initial_Track_2),
    sum(Initial_Track_3),
    sum(Current_Track_1),
    sum(Current_Track_1_Plus),
    sum(Current_Track_2),
    sum(Current_Track_3)
  ) %>%
  gather(key = "Type", value = "Count") %>%
  separate("Type", into = c("A", "Type", "C", "Track","E")) %>%
  select(-A, -C) %>%
  unite("Track", c("Track", "E"), sep = " ")
  
ggplot(tracks, aes(x = Track, y = Count, fill = Type)) + 
  geom_col(position = "dodge") +
  labs(
    title = "Number of ACOs by Track",
    subtitle = "between Initial and Current enrollment",
    y = "# of ACOs"
  )
```
```{r}
# Generated Total Savings / Losses

aco %>% select(GenSaveLoss) %>%
  mutate(Amount = "Zero") %>%
  mutate(Amount = ifelse(GenSaveLoss > 0, "Positive", Amount)) %>%
  mutate(Amount = ifelse(GenSaveLoss < 0, "Negative", Amount)) %>%
  count(Amount)

ggplot(aco, aes(x=GenSaveLoss, fill = Agree_Type)) + 
  geom_density(alpha = 0.4) +
  scale_x_continuous(labels = scales::comma) +
  labs(
    title = "Distribution of Generated Losses and Savings (in Dollars)",
    x = "Generated Losses and Savings"
  )

ggplot(aco %>% filter(GenSaveLoss != 0), 
       aes(x=GenSaveLoss, fill = Agree_Type)) + 
  geom_density(alpha = 0.4) +
  scale_x_continuous(labels = scales::comma) +
  labs(
    title = "Distribution of non-Zero Generated Losses and Savings (in Dollars)",
    x = "Generated Losses and Savings"
  ) + geom_vline(aes(xintercept = mean(GenSaveLoss, color = Agree_Type)))
  
```


There are 4 types of beneficiaries that are categorized to account for risk adjustments:

1. ESRD (End Stage Renal Disease)
    Eligible for Medicare as a result of End Stage Renal Disease
2. Disabled
    Eligible for Medicare as a result of disability
3. Aged/Dual
    Eligible for Medicare by age (65+) and eligible for Medicaid
4. Aged/Non-Dual
    Eligible only for Medicare by age (65+)


