---
title: "Extract Transform Load - ACO"
author: "Karen Jiang"
date: "2/26/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(readr)
library(readxl)
library(gt)
```

## Exploratory Data Analysis - Part 1

We are examining 3 datasets from CMS.

* **ACO general information**
* Beneficiaries information by ACO
* Fee-for-Service Expenditures by ACO

To better understand how MSSP scoring and classifications for ACOs, information can be found in the  [MSSP Shared Savings and Losses Assignement and Methodology ]("https://www.cms.gov/Medicare/Medicare-Fee-for-Service-Payment/sharedsavingsprogram/Downloads/Shared-Savings-Losses-Assignment-Spec-V7.pdf").



```{r read_data, message=FALSE}

# Read in .csv data

aco <- 
  read.csv("data/2018_Shared_Savings_Program__SSP__Accountable_Care_Organizations__ACO__PUF.csv") %>%
  select(-contains("N_"), -contains("P_")) 

# Read in .xlsx data dictionary

aco_dictionary <-
  read_excel("data-dictionary/Dictionary.ACO.SSP.PUF.Y2018.xlsx",
             sheet = "ACO PUF Dictionary 2018",
             skip = 16,
             col_names = TRUE)

```

```{r functions}
# Function to look in the dictionary and find the definition of a column. 

find_definition_aco <- function(variable, col = c("short", "long", "type")) {
  col <- match.arg(col)
  aco_dictionary %>%
    filter(Variable == variable) %>%
    pull(
      switch(col,
         short = "Short Description",
         long = "Long Description",
         type = "Format/Values"))
}
```

### Understanding the ACO dataset

I wanted to look at how ACOs were distributed in terms of the ACO Track Type. For further information on the Track Types, [read here](http://www.physiciansadvocacyinstitute.org/Portals/0/assets/docs/Advanced-APM-Pathway/Medicare%20Shared%20Savings%20Program%20Overview.pdf).

- Track 1 : One-sided risk
- Track 1+ : Two-sided risk
- Track 2 : Two-sided risk
- Track 3 : Two-sided risk


```{r aco, include = FALSE}

aco %>%
  count(Agree_Type)

tracks <- aco %>%
  summarise(
    sum(Initial_Track_1),
    sum(Initial_Track_1_Plus),
    sum(Initial_Track_2),
    sum(Initial_Track_3),
    sum(Current_Track_1),
    sum(Current_Track_1_Plus),
    sum(Current_Track_2),
    sum(Current_Track_3)
  ) %>%
  gather(key = "Type", value = "Count") %>%
  separate("Type", into = c("A", "Type", "C", "Track","E")) %>%
  select(-A, -C) %>%
  unite("Track", c("Track", "E"), sep = " ")
  
ggplot(tracks, aes(x = Track, y = Count, fill = Type)) + 
  geom_col(position = "dodge") +
  labs(
    title = "Number of ACOs by Track",
    subtitle = "between Initial and Current enrollment",
    y = "# of ACOs"
  )
```


```{r}
# Generated Total Savings / Losses

aco %>% select(GenSaveLoss) %>%
  mutate(Amount = "Zero") %>%
  mutate(Amount = ifelse(GenSaveLoss > 0, "Positive", Amount)) %>%
  mutate(Amount = ifelse(GenSaveLoss < 0, "Negative", Amount)) %>%
  count(Amount)

ggplot(aco, aes(x=GenSaveLoss, fill = Agree_Type)) + 
  geom_density(alpha = 0.4) +
  scale_x_continuous(labels = scales::comma) +
  labs(
    title = "Distribution of Generated Losses and Savings (in Dollars)",
    x = "Generated Losses and Savings"
  )

ggplot(aco %>% filter(GenSaveLoss != 0), 
       aes(x=GenSaveLoss, fill = Agree_Type)) + 
  geom_density(alpha = 0.4) +
  scale_x_continuous(labels = scales::comma) +
  labs(
    title = "Distribution of non-Zero Generated Losses and Savings (in Dollars)",
    x = "Generated Losses and Savings"
  ) + geom_vline(aes(xintercept = mean(GenSaveLoss, color = Agree_Type)))
  
```


There are 4 types of beneficiaries that are categorized to account for risk adjustments:

1. ESRD (End Stage Renal Disease)
    Eligible for Medicare as a result of End Stage Renal Disease
2. Disabled
    Eligible for Medicare as a result of disability
3. Aged/Dual
    Eligible for Medicare by age (65+) and eligible for Medicaid
4. Aged/Non-Dual
    Eligible only for Medicare by age (65+)


