---
title: "Extract Transform Load - Beneficiaries"
author: "Karen Jiang"
date: "3/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidycensus)
library(readr)
library(readxl)
library(tidyverse)
library(ggthemes)
library(stringr)
library(janitor)
library(sf)
```

## Exploratory Data Analysis - Part 2

* ACO general information
* **Beneficiaries information by ACO**
* **Fee-for-Service Expenditures by County**

To better understand how MSSP scoring and classifications for ACOs, information can be found in the  [MSSP Shared Savings and Losses Assignement and Methodology ]("https://www.cms.gov/Medicare/Medicare-Fee-for-Service-Payment/sharedsavingsprogram/Downloads/Shared-Savings-Losses-Assignment-Spec-V7.pdf").


Our goal with the data is to understand:

- How does ACOs saturation by state?
- Within states, how are ACOs distributed by county?
- How are person-years contributed to each state?
- What are the risk scores by county?
- Is there any correlation between risk scores and ACO saturation?

```{r read_data, include=FALSE}

# GEODATA

# Variable "B00001_001" is a county population estimate.

# census_api_key("6c3bf75163fa1ef1eb5aff46febb50bfb040b127")

geo_data <- get_acs(geography = "county",
                  variables = "B19013_001", 
                  year = 2018,
                  geometry = TRUE) %>%
  separate(NAME, into = c("County", "State"), ", ") %>%
  mutate(County = str_remove(string = County, pattern = "County"))


# BENEFICIARIES DATA

beneficiaries <- 
  read_csv("data/2018_Number_of_ACO_Assigned_Beneficiaries_by_County_PUF.csv") %>%
  clean_names() %>%
  mutate(geoid = paste(state_id, county_id, sep=""))

beneficiaries_dictionary <-
  read_excel("data-dictionary/Dictionary.ACO.SSP.COUNTY.ASSIGNED.PUF.2018.xlsx",
             sheet = "Data Dictionary",
             col_names = TRUE)

# COUNTY FEE FOR SERVICE DATA

ffs <- 
  read_csv("data/2018_County-level_FFS_Data_for_Shared_Savings_Program_Benchmark_PUF.csv") %>%
  clean_names() %>%
  mutate(state_id = str_pad(state_id, width = 2,pad = "0"),
         county_id = str_pad(county_id, width = 3, pad = "0")) %>%
  mutate(geoid = paste(state_id, county_id, sep=""))

ffs_dictionary <-
  read_excel("data-dictionary/Dictionary.ACO.SSP.COUNTYRATE.PUF.PY19.2018.xlsx",
             sheet = "Data Dictionary",
             col_names = TRUE)

```

```{r functions, include=FALSE}

# Helper functions to quickly call on the variable definitions.

find_definition_beneficiaries <- function(variable, col = c("short", "long", "type")) {
  col <- match.arg(col)
  beneficiaries_dictionary %>%
    filter(`Variable Name` == variable) %>%
    pull(
      switch(col,
         short = "Short Description",
         long = "Long Description",
         type = "Missing or Suppressed Values"))
}


find_definition_ffs <- function(variable, col = c("short","long","type")) {
  col <- match.arg(col)
  ffs_dictionary %>%
    filter(`Variable Name` == variable) %>%
    pull(
      switch(col,
             short = "Short Description",
             long = "Long Description",
             type = "Missing or Suppressed Values"))
}

# Example: 
# find_definition_ffs("County_ID", "long")
```


### Understanding the Beneficiaries dataset

#### Saturation

```{r}
unique_acos <- length(unique(beneficiaries$`ACO ID`))
```
The goal of the MSSP program is to encourage greater adoption of ACOs as alternative payment models. Here, we will explore the saturation of ACOs in the US by state and by county. There are `r unique_acos` unique ACOs in this dataset. 

```{r aco_state}
num_acos_by_state <- beneficiaries %>% 
  group_by(state_name) %>%
  count(aco_id) %>%
  count(state_name)

ggplot(num_acos_by_state, aes(y= reorder(state_name, n), x=n)) + 
  geom_text(aes(label = n, size = "12")) + 
  theme_classic() +
  theme(
    axis.ticks = element_blank(),
    aspect.ratio = 2.5,
    axis.line.x.bottom = element_blank(),
    axis.text.x.bottom = element_blank(),
    legend.position = "none"
  ) +
  labs(
    x = "Number of ACOs",
    y = "US States and Territories"
  )

```
If we look within a State, say Massachusetts, ACOs may also be unevenly distributed within between counties.

```{r geo_state}

ffs %>%
  group_by(state_name) %>%
  count(state_name)

cali_data <- geo_data %>% right_join(ffs, by = c("State" = "state_name", "County" = "county_name")) %>% filter(State == "California") %>%
  mutate(aged_non_dual_person_years = as.numeric(aged_non_dual_person_years))

cali_data %>%
  ggplot() +
  geom_sf() +
  scale_fill_viridis_c(option = "plasma",
                       direction = -1)


nebraska <- income %>%
  filter(State == "Nebraska")

ggplot(nebraska, aes(fill = estimate)) + geom_sf()

# geo_ffs %>% 
#   filter(state_name == "Alaska") %>%
#   ggplot(aes(geometry = geometry, fill = b00001_001e)) +
#   geom_sf()
```


We can first choose to look at the total number of assigned beneficiaries per county. The values for this dataset are: `r find_definition_beneficiaries("Tot_AB", "type")`

What proportion of counties have less than 10 Total Assigned Beneficiaries? 
```{r}
beneficiaries %>%
  filter(total_assigned_beneficiaries == "*") %>%
  length()
```

